Hourly Assimilation of Different Sources of Observations Including Satellite Radiances in a Mesoscale Convective System Case During RELAMPAGO campaign

### 2022-08-24 analysis/paper/paper_ana_20181122.Rmd
Chau we

  In this paper, we evaluate the impact of assimilating high-resolution surface networks and satellite observations using the WRF-GSI-LETKF over central and north eastern Argentina where the surface and upper air observing networks are relatively coarse. We conducted a case study corresponding to a huge mesoscale convective system (MCS) that developed during November 22, 2018. The accumulated precipitation associated with this MCS was quite high, exceeding 200 mm over northern Argentina and Paraguay. The MCS developed during the Intense Observing Period (IOP) of the Remote sensing of Electrification, Lightning, And Mesoscale/microscale Processes with Adaptive Ground Observations (RELAMPAGO) field campaign. We used the GSI-4DLETKF data assimilation package to produce analyses by assimilating observations every hour with 10-km horizontal grid spacing and a 60-member multiphysics ensemble. We conducted four assimilation experiments using different sets of observations: CONV, consisting of conventional observations from NCEP’s prepBUFR files; AWS, combining CONV and dense automatic surface weather station networks, SATWND, combining AWS with satellite-derived winds, and RAD, including SATWND; and satellite radiances from different microwave and infrared sensors. We found that the assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed in this case study. The assimilation of radiance observations produces a better development of the convection mainly during the mature state of the MCS leading to an increase in the accumulated precipitation. We also ran ensemble forecasts initialized from each experiment and evaluated their skill to predict precipitation. We found that the hourly assimilation of the observations in AWS, SATWND, and RAD helped to improve the precipitation forecast.

@@

classoption: preprint, 3p, authoryear

@@

For DA methods to be successful, observing networks with sufficient temporal and spatial resolution capable of capturing mesoscale variability should be used [@gustafsson2018]. Assimilating information on temperature, moisture and wind in the planetary boundary layer improves mesoscale model initialization, and several authors have reported the resultant beneficial impacts upon the planetary boundary layer structure and the location and timing of precipitating systems (e.g. @wheatley2010, @ha2014, @chang2017, @bae2022, @banos2021, @maejima2019, and @chen2016).

Particularly relevant for regional mesoscale DA systems in the region of interest is that South America is characterized by a limited number of conventional observations (i.e., radiosondes, surface weather stations) and operational networks are not dense enough to capture mesoscale details. In this context, analyzing the potential impact of non-conventional sources of observations is essential to improve mesoscale numerical weather prediction (NWP) over South America using DA. There have been only a few published efforts on regional mesoscale DA, but they have all shown promising results [e.g. @dillon2016; @dillon2021; @goncalvesdegoncalves2015]. In particular @dillon2021 assimilated high resolution surface weather station networks, GOES-16 atmospheric motion vectors (AMVs), and satellite temperature and moisture retrievals over central Argentina with positive impacts. Similarly to @gasperoni2018, @dillon2021 included private weather station networks which are not incorporated in the operational analysis. However, the impact of different observation types on the analysis quality has not been addressed.

The impact of non-conventional high spatial and temporal resolution observations, such as AMVs, has been investigated in the context of regional mesoscale DA. Many studies have focused on the impact of these observations on the prediction of tropical storms (e.g., @wu2014, @cherubini2006, and @sawada2019, and many others). Most of these studies reported an overall positive impact of the assimilation of AMVs for this type of storm. However, some works indicated mixed impacts (e.g. @sawada2019 reported an improvement in the forecast of the track of the storm but a degradation in the forecast intensity). As stated in @zhao2021 and @zhao2021a, the impact of assimilating these data on high impact weather events associated with mid-latitude deep convection over land has received relatively less attention. \citet{zhao2021, zhao2021a} assimilated GOES-16 AMVs into a storm-scale three-dimensional variational DA system during three high impact weather events. They reported positive impacts of AMVs on the characterization of the storm environment and improved short range precipitation forecasts. @otsuka2015 and @mallick2020 found a slight improvement in the short-range precipitation forecast due to the storm-scale assimilation of high frequency AMVs.

@@

The main objective of this work is thus to contribute to the quantification and comparison of the impact of high resolution automatic weather stations, AMVs, and clear-sky satellite radiances, into a mesoscale, frequently-updated ensemble-based regional DA system. This is particularly important in the efforts to improve mesoscale numerical weather prediction (NWP) over South America where the conventional observation network is rather sparse and other sources of information could potentially fill certain gaps. In particular, we will focus on the impact in the context of a mid-latitude mesoscale convective system.

To reach this goal, we conduct several DA experiments for a case study of a large Mesoscale Convective System (MCS) that developed over Southern South America during Nov 22-23, 2018 during the intense observation period (IOP) of the RELAMPAGO field campaign.

@@

(ref:case) ERA5 Reanalysis of pressure (hPa) at sea level (black contours), 1000-500 hPa thickness (red dashed contours) and convective available potential energy (shaded) for 00 and 12 UTC Nov 22 and 00 UTC Nov 23.

@@

The forecast model uses the non-hydrostatic Advanced Research version of Weather Research and Forecasting (WRF-ARW V3.9.1, @skamarock2008). 
The horizontal grid spacing is 10 km (150 x 200 grid points) in the horizontal and 37 levels in the vertical with the top of the model at 50 hPa. 
The initial and boundary conditions are provided by the Global Forecast System (GFS) analysis (0.25$^{\circ}$ horizontal grid spacing and 6-hour temporal resolution; @cisl_rda_ds084.1). “In this case a single nesting approach is used since the resolution gap between the driving model and the regional model is not too large (0.25$^{\circ}$ or 25 km aproximately to 10 km). This approach is also based in recent studies which suggest that using multiple nested domains do not necessarily lead to improved precipitation forecasts in regional domains, particularly in areas of complex terrain (e.g. @liang2019, @beck2004). The domain covers the area indicated in Figure \@ref(fig:dominio) to capture the development of the MCS during the simulated period.

@@

We use a 60-member ensemble where the initial ensemble mean and the mean boundary conditions are taken from the GFS deterministic analysis. A set of 60 perturbations are  randomly generated to perturb the initial state as well as the  boundary conditions during the length of the experiment. Perturbing the boundary conditions helps to reduce the impact of errors in the driving global model and helps to keep a larger ensemble spread throughout the domain and during the length of the experiment [@ouaraini2015]. The perturbations are generated as scaled differences between two random atmospheric states obtained from the Climate Forecast System Reanalysis (CFSR) data with 0.5$^{\circ}$ horizontal grid spacing with a smooth time evolution as in @necker2020 and @maldonado2021. In this way, we preserved the nearly hydrostatic and geostrophic equilibrium of larger scales. The random perturbations are the same across different experiments to ensure that the differences between experiments are only related to changes in the number and type of assimilated observations.

We used multi-physics scheme to better represent the uncertainty in the model formulation within the DA system. We generated 9 different model configurations consisting of the combination of 3 moist convection schemes (Kain–Fritsch [@kain2004], Grell–Freitas [@grell2013], and Betts–Miller–Janjic [@janjic1994]) and 3 planetary boundary layer schemes (Yonsei University Scheme [@hong2006], Mellor–Yamada–Janjic Scheme [@janjic1994], and Mellor–Yamada Nakanishi Niino [@nakanishi2009]). The distribution of these schemes among the 60 ensemble members is outlined in Table \@ref(tab:miembros-desc). The multi-physics approach is also introduced in order to represent the uncertainty associated with the more relevamnt physical processes that are not resolved by the model. All ensemble members use the same land-surface model (Noah-MP, @chen2001), microphysics (WRF single-moment 6–class scheme [@hong2006a]), and radiation processes (RRTMG shortwave and longwave scheme [@iacono2008]) parameterizations.

```{r miembros-desc}
fisica %>% 
  copy() %>% 
  .[, c("Cumulus", "PBL") := tstrsplit(fisica, split = "-")] %>% 
  .[, .(mem = paste(as.numeric(mem), collapse = ", ")), by = .(PBL, Cumulus)] %>% 
  dcast(Cumulus ~ PBL, value.var  = "mem") %>% 
  kable(booktabs = TRUE, caption = "Generation of the 60-member RRR multi-physics ensemble as a combination of Cumulus and PBL parameterizations, and GEFS members for boundary conditions.", 
        align = "cccc") %>% 
  add_header_above(c(" " = 1, PBL = 3)) %>% 
  kable_classic_2(full_width = FALSE) %>% 
  kable_styling(font_size = 6,
                position = "center") %>% 
  # column_spec(1, width = "1em") %>% 
  column_spec(2:4, width = "7em")
```

To reduce the effect of spurious correlations in the estimation of error covariances, we use a horizontal localization radius of 180 km and a vertical localization radius of 0.4 (in log pressure coordinates) as in @dillon2021 for all types of observations. 
A relaxation-to-prior spread inflation [@whitaker2012] is applied with an inflation parameter $\alpha=0.9$ following @maldonado2020 to mitigate the impact of sampling errors and to consider model errors not accounted for by the multi-model ensemble approach.

@@

The conventional observations used are part of the Global Data Assimilation System (GDAS) data stream. We assimilated conventional observations included in the Binary Universal Form for Representation of Meteorological Data (PREPBUFR) files generated at the National Centers for Environmental Prediction (NCEP). These consist of surface observations from 117 Conventional Surface Weather Stations (CSWS), ships, and upper-air observations from 13 radiosondes sites and aircraft. The orange triangles in Figure \@ref(fig:dominio)a indicate the location of the surface stations included in this experiment. The frequency of these observations varied between 1 hour for surface stations and 12/24 hours for radiosondes. Wind surface observations over oceans (ASCATW) come from scatterometers and are also included in the PREPBUFR files.

@@

We also assimilate data from 866 Automatic Weather Stations (AWS) that are part of 17 public and private surface networks over Southern South America. The dataset used in this study has been obtained from the RELAMPAGO Data Set repository [@garcia2019]. These stations are indicated as green squares in Figure \@ref(fig:dominio)a. They have higher spatial coverage than the CSWS and a sampling frequency of 10 minutes in most cases. All stations measure temperature, but only 395 stations provide humidity, 422 provide pressure, and 605 provide wind information. 
Observation errors used to assimilate these observations are the same as for the CSWS (see Table \@ref(tab:table-obs)).

@@

To evaluate the performance of the ensemble-based DA system presented in this article, we use the following observational datasets:

@@

To investigate the impact of different observations upon the analysis, four DA experiments were performed using different observation sets (Table \@ref(tab:table-exp)). The CONV experiment uses only conventional observations from PREPBUFR. In a second experiment, referred to as AWS, we assimilate all the observations included in CONV plus the 10-minute frequency surface observations from AWS. In the third experiment, referred to as SATWND, we assimilate all the observations of the AWS experiment and the satellite-derived winds. Finally, a fourth experiment referred to as RAD assimilates all available clear-sky radiances from sensors onboard polar orbiting satellites as described in section \@ref(sat).

@@

To evaluate the statistical consistency of the uncertainty quantification in the ensemble system we use the Reduced Centered Random Variable (RCRV, @candille2007) which is defined as:

@@

where $x_o$ is the assimilated observation and its error $\sigma_o$, the ensemble mean of the analysis in observational space $m$, and the standard deviation $\sigma$ of the ensemble.  The $RCRV$ is the ratio between the distance between the observations and the forecast and its expected standard deviation assuming the statistical independence between the forecast error (estimated from the ensemble spread) and the observation error. The average of $RCRV$ computed over all the realizations represents the bias of the ensemble mean with respect to the observations normalized by the estimated uncertainty:

@@

For the comparison of the first-guess precipitation with the IMERG precipitation estimates, we compute the Fractions Skill Score (FSS, @roberts2008) for different neighborhood length scales and thresholds:

@@

where $P_{oi}$ is the fraction of grid points in the $i-th$ sampling area in which the observed accumulated precipitation is greater than a specified threshold. Following @roberts2020, $P_{xi}$ is calculated from the ensemble probability precipitation over the same threshold in each grid point by averaging over the $i-th$ sampling area. 
The FSS was computed from the accumulated precipitation over 6 hr periods by adding the 1-hr accumulated precipitation forecasts over 6 consecutive assimilation cycles.

@@

We performed all the experiments at the National Center for Atmospheric Research (NCAR) supercomputer Cheyenne [@Cheyenne2019]. All the analyses in this paper were performed using the R programming language [@rcoreteam2020], using data.table [@dowle2020] and metR [@campitelli2020] packages.
All graphics are made using ggplot2 [@wickham2009] and the paper was rendered using knitr and rmarkdown [@xie2015; @allaire2019].

@@

To investigate the ability of the first-guess ensemble mean to fit the observations taking into account the uncertainties of the forecast and the observations, we calculated the $mean RCRV$ and the $sd RCRV$ for the RAD experiment. As this experiment assimilates all types of observations used in this work, it is possible to analyze the consistency of the ensemble by comparing it with each type of observation. Figure \@ref(fig:rcrv-sfc) shows the $sd RCRV$ for surface observations box-averaged to a 2.5° grid. The $sd RCRV$ for wind observations (Figure \@ref(fig:rcrv-sfc)a) is close to 1 suggesting a good agreement between the ensemble spread, the forecast error, and the observation error. For the temperature (Figure \@ref(fig:rcrv-sfc)b), the results are similar except that for some areas in the west of the domain the $sd RCRV$ can be as high as 4.5. These higher values of $sdRCRV$ can be associated with systematic errors arising from heigh differences between the model surface and the observations. Small scale circulations associated with the complex terrain and not well resolved by the model can also contribute to increase the distance between the forecast and the observations. These aspects are usually not captured by the ensemble spread unless a well tuned space dependent inflation scheme is used thus leading to greater sdRCRV values.

Figure \@ref(fig:rcrv-profile) shows the mean and standard deviation of the RCRV for the upper-air observations. Figures \@ref(fig:rcrv-profile)a-b show the RCRV statistics for soundings (ADPUPA) and aircraft (AIRCAR and AIRCFT). Both ADPUPA and AIRCFT show a generally good agreement between the ensemble spread and the observation error. As sounding observations and their associated errors are known to be reliable, this result indicates that the ensemble has an appropriate spread. AIRCAR presents an irregular profile with $sd RCRV$ values that suggest that the error for this type of observation is overestimated. ADPUPA and AIRCAR present a $mean RCRV$ profile near zero at middle and upper levels. At low levels, the $mean RCRV$ profile is positive, showing a cold bias present in the model, a characteristic already studied in @ruiz2010 and @dillon2021.

Satellite-derived winds observations varies in number depending on the satellite and the level. In Figure \@ref(fig:rcrv-profile)c we included the $RCRV$ calculated at least with 100 observations for each satellite and level. At low levels, where there are not many observations available, the profiles of $mean RCRV$ and $sd RCRV$ show a larger departure from the expected behavior with a negative bias, and a possible overestimation of the observation error. Wind estimations derived from water vapor channels are abundant above 500 hPa where their bias is close to zero. The only exception are the EUMETSAT observations which contribute very little in the region.

@@

In this section, we present the impact of assimilating different observation types on variables which are particularly relevant for the occurrence of deep moist convection. The analysis is performed over a smaller domain (red box in Figure \@ref(fig:dominio)a) to focus on the region most directly affected by the MCS. Figures \@ref(fig:TQ-diff)a-c show the analysis difference between experiments in the spatially averaged vertical profile of temperature. By averaging the differences between two experiments we can isolate the systematic impact produced by different observing systems on the analyzed state. During the first day, the assimilation of AWS observations results in a colder PBL. This cooling effect has a clear diurnal cycle, being stronger during nighttime (Figure \@ref(fig:TQ-diff)a). During the second day of the experiment, the impact of AWS observations extends into the middle and upper troposphere coinciding with the mature stage of the MCS. The warm difference shown in AWS-CONV between 500 and 200 hPa is produced by the development of stronger convection in AWS compared to CONV. This is a good example of how low-level information provided by surface weather stations can rapidly spread into the troposphere in the presence of deep moist convection. Although the mid-to-upper circulation can have an important impact on the organization and evolution of the MCS over the region, the satellite-derived winds did not have an appreciable impact on the mean temperature and humidity (Figure \@ref(fig:TQ-diff)b-e), possibly due to the large observation errors used for the assimilation.
During the first day of the experiment, the assimilation of radiances produces a warming effect in the PBL which partially compensates for the cooling effect of AWS observations (Figure \@ref(fig:TQ-diff)c). No clear systematic impact is found above the PBL during this period. During the second day, the impact of radiance observations is found through the troposphere with a distribution that is similar to the impact found in the AWS experiment but with the opposite sign.

Comparing the specific humidity in the experiments (Figures \@ref(fig:TQ-diff)d-f), the impact of assimilating AWS with fine spatial and temporal resolution is most substantial at low levels (Figure \@ref(fig:TQ-diff)d). The PBL in the AWS experiment is consistently moister than in the CONV experiment particularly at nighttime. The increase in low-level moisture by a denser surface network is consistent with previously reported dry biases in the WRF model over the region [@casaretto2022, @matsudo2021, @ruiz2010]. We found that the moistening of the PBL is mainly driven by the covariance between temperature and specific humidity within the PBL. In our experiment and over the center of the domain, this covariance remains negative, increasing low-level moisture as the observations introduce negative temperature corrections. As for the temperature, the systematic impact of satellite-derived winds on moisture is small (Figure \@ref(fig:TQ-diff)e). Figure \@ref(fig:TQ-diff)f shows that radiances reduce low-middle level moisture during the first day of the experiment. The drying effect extends to lower-middle levels during the second day of the experiment coinciding with the development of the MCS between 00 and 12 UTC Nov 22.

@@

IOP 7 (Figures \@ref(fig:soundings)a-d) provides a good characterization of the pre-convective environment during the first day of our experiments. The area where the observations were taken was characterized by mostly clear skies and a low-level northerly flow associated with warm and moist advection. In general, the experiments show a similar RMSE and bias for all the variables. AWS observations were able to reduce the RMSE for temperature and dew point temperature in the PBL and reduce a small dry bias. However, in this region (Figure \@ref(fig:dominio)b) and for this period, AWS increments (Figure \@ref(fig:UV-diff)d) degradated the zonal wind between 7 and 12 km increasing the bias and RMSE (Figure \@ref(fig:soundings)c).

>>>>

  This paper evaluates the impact of assimilating high-resolution surface networks and satellite observations using the WRF-GSI-LETKF over central and north eastern Argentina where the surface and upper air observing networks are relatively coarse. We conducted a case study corresponding to a huge mesoscale convective system (MCS) that developed during November 22, 2018. The accumulated precipitation associated with this MCS was quite high, exceeding 200 mm over northern Argentina and Paraguay. The MCS developed during the Intense Observing Period (IOP) of the Remote sensing of Electrification, Lightning, And Mesoscale/microscale Processes with Adaptive Ground Observations (RELAMPAGO) field campaign. The GSI-4DLETKF data assimilation package is used to produce analyses by assimilating observations every hour with 10-km horizontal grid spacing and a 60-member multiphysics ensemble. Four assimilation experiments are conducted using different sets of observations: CONV, consisting of conventional observations from NCEP’s prepBUFR files; AWS, combining CONV and dense automatic surface weather station networks, SATWND, combining AWS with satellite-derived winds, and RAD, including SATWND; and satellite radiances from different microwave and infrared sensors. The assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed in this case study. The assimilation of radiance observations produces a better development of the convection mainly during the mature state of the MCS leading to an increase in the accumulated precipitation. Ensemble forecasts initialized from each experiment were also similated to evaluated their skill to predict precipitation. The hourly assimilation of the observations in AWS, SATWND, and RAD helped to improve the precipitation forecast.

@@

classoption: preprint, 5p, authoryear, twocolumn

@@

For DA methods to be successful, observing networks with sufficient temporal and spatial resolution capable of capturing mesoscale variability should be used [@gustafsson2018]. Assimilating information on temperature, moisture and wind in the planetary boundary layer (PBL) improves mesoscale model initialization, and several authors have reported the resultant beneficial impacts upon the PBL structure and the location and timing of precipitating systems (e.g. @wheatley2010, @ha2014, @chang2017, @bae2022, @banos2021, @maejima2019, and @chen2016).

Particularly relevant for regional mesoscale DA systems in the region of interest is that South America is characterized by a limited number of conventional observations (i.e., radiosondes, surface weather stations) and operational networks that are not dense enough to capture mesoscale details. In this context, analyzing the potential impact of non-conventional sources of observations is essential to improve mesoscale numerical weather prediction (NWP) over South America using DA. There have been only a few published efforts on regional mesoscale DA, but they have all shown promising results [e.g. @dillon2016; @dillon2021; @goncalvesdegoncalves2015]. In particular @dillon2021 assimilated high resolution surface weather station networks, GOES-16 atmospheric motion vectors (AMVs), and satellite temperature and moisture retrievals over central Argentina with positive impacts. Similarly to @gasperoni2018, @dillon2021 included private weather station networks which are not incorporated in the operational analysis. However, the impact of different observation types on the analysis quality has not been addressed.

The impact of non-conventional high spatial and temporal resolution observations, such as AMVs, has been investigated in the context of regional mesoscale DA. Many studies have focused on the impact of these observations on the prediction of tropical storms (e.g., @wu2014, @cherubini2006, and @sawada2019, and many others). Most of these studies reported an overall positive impact of the assimilation of AMVs for this type of storm. However, some works indicated mixed impacts (e.g. @sawada2019 reported an improvement in the forecast of the track of the storm but a degradation in the forecast intensity). As stated in \citet{zhao2021, zhao2021a}, the impact of assimilating these data on high impact weather events associated with mid-latitude deep convection over land has received relatively less attention. \citet{zhao2021, zhao2021a} assimilated GOES-16 AMVs into a storm-scale three-dimensional variational DA system during three high impact weather events. They reported positive impacts of AMVs on the characterization of the storm environment and improved short range precipitation forecasts. @otsuka2015 and @mallick2020 found a slight improvement in the short-range precipitation forecast due to the storm-scale assimilation of high frequency AMVs.

@@

The main objective of this work is thus to contribute to the quantification and comparison of the impact of high resolution automatic weather stations, AMVs, and clear-sky satellite radiances, into a mesoscale, frequently-updated ensemble-based regional DA system. This is particularly important in the efforts to improve mesoscale numerical weather prediction (NWP) over South America where the conventional observation network is rather sparse and other sources of information could potentially fill certain gaps. In particular, this paper focus on the impact in the context of a mid-latitude mesoscale convective system. To reach this goal, several DA experiments are conducted for a case study of a large Mesoscale Convective System (MCS) that developed over Southern South America during Nov 22-23, 2018 during the intense observation period (IOP) of the RELAMPAGO field campaign.

@@

(ref:case) ERA5 Reanalysis of pressure (hPa) at sea level (black contours), 1000-500 hPa thickness (red dashed contours) and convective available potential energy (shaded) for a) 00 and b) 12 UTC Nov 22 and c) 00 UTC Nov 23.

@@

The forecast model uses the non-hydrostatic Advanced Research version of Weather Research and Forecasting (WRF-ARW V3.9.1, @skamarock2008). 
The horizontal grid spacing is 10 km (150 x 200 grid points) in the horizontal and 37 levels in the vertical with the top of the model at 50 hPa. 
The initial and boundary conditions are provided by the Global Forecast System (GFS) analysis (0.25$^{\circ}$ horizontal grid spacing and 6-hour temporal resolution; @cisl_rda_ds084.1). In this case a single nesting approach is used since the resolution gap between the driving model and the regional model is not too large (0.25$^{\circ}$ or 25 km approximately to 10 km). This approach is also based in recent studies which suggest that using multiple nested domains do not necessarily lead to improved precipitation forecasts in regional domains, particularly in areas of complex terrain (e.g. @liang2019, @beck2004). The domain covers the area indicated in Figure \@ref(fig:dominio) to capture the development of the MCS during the simulated period.

@@

A 60-member ensemble is used where the initial ensemble mean and the mean boundary conditions are taken from the GFS deterministic analysis. A set of 60 perturbations are randomly generated to perturb the initial state as well as the boundary conditions during the length of the experiment. Perturbing the boundary conditions helps to reduce the impact of errors in the driving global model and helps to keep a larger ensemble spread throughout the domain and during the length of the experiment [@ouaraini2015]. The perturbations are generated as scaled differences between two random atmospheric states obtained from the Climate Forecast System Reanalysis (CFSR) data with 0.5$^{\circ}$ horizontal grid spacing with a smooth time evolution as in @necker2020 and @maldonado2021. In this way, the nearly hydrostatic and geostrophic equilibrium of larger scales is preserved. The random perturbations used are the same across experiments to ensure that the differences between experiments are only related to changes in the number and type of assimilated observations.

A multi-physics scheme is used to better represent the uncertainty in the model formulation within the DA system. 9 different model configurations are generated consisting of the combination of 3 moist convection schemes (Kain–Fritsch [@kain2004], Grell–Freitas [@grell2013], and Betts–Miller–Janjic [@janjic1994]) and 3 planetary boundary layer schemes (Yonsei University Scheme [@hong2006], Mellor–Yamada–Janjic Scheme [@janjic1994], and Mellor–Yamada Nakanishi Niino [@nakanishi2009]). The distribution of these schemes among the 60 ensemble members is outlined in Table \@ref(tab:miembros-desc). The multi-physics approach is also introduced in order to represent the uncertainty associated with the more relevant physical processes that are not resolved by the model. All ensemble members use the same land-surface model (Noah-MP, @chen2001), microphysics (WRF single-moment 6–class scheme [@hong2006a]), and radiation processes (RRTMG shortwave and longwave scheme [@iacono2008]) parameterizations.

```{r miembros-desc}
fisica %>% 
  copy() %>% 
  .[, c("Cumulus", "PBL") := tstrsplit(fisica, split = "-")] %>% 
  .[, .(mem = paste(as.numeric(mem), collapse = ", ")), by = .(PBL, Cumulus)] %>% 
  dcast(Cumulus ~ PBL, value.var  = "mem") %>% 
  kable(booktabs = TRUE, caption = "Generation of the 60-member multi-physics ensemble as a combination of Cumulus and PBL parameterizations.", 
        align = "cccc") %>% 
  add_header_above(c(" " = 1, PBL = 3)) %>% 
  kable_classic_2(full_width = FALSE) %>% 
  kable_styling(font_size = 6,
                position = "center") %>% 
  # column_spec(1, width = "1em") %>% 
  column_spec(2:4, width = "7em")
```

To reduce the effect of spurious correlations in the estimation of error covariances, a horizontal localization radius of 180 km and a vertical localization radius of 0.4 (in log pressure coordinates) is used as in @dillon2021 for all types of observations. 
A relaxation-to-prior spread inflation [@whitaker2012] is applied with an inflation parameter $\alpha=0.9$ following @maldonado2020 to mitigate the impact of sampling errors and to consider model errors not accounted for by the multi-model ensemble approach.

@@

The conventional observations used are part of the Global Data Assimilation System (GDAS) data stream. Conventional observations included in the Binary Universal Form for Representation of Meteorological Data (PREPBUFR) files generated at the National Centers for Environmental Prediction (NCEP) are assimilated. These consist of surface observations from 117 Conventional Surface Weather Stations (CSWS), ships, and upper-air observations from 13 radiosondes sites and aircraft. The orange triangles in Figure \@ref(fig:dominio)a indicate the location of the surface stations included in this experiment. The frequency of these observations varied between 1 hour for surface stations and 12/24 hours for radiosondes. Wind surface observations over oceans (ASCATW) come from scatterometers and are also included in the PREPBUFR files.

@@

Data from 866 Automatic Weather Stations (AWS) that are part of 17 public and private surface networks over Southern South America are also assimilated. The dataset used in this study has been obtained from the RELAMPAGO Data Set repository [@garcia2019]. These stations are indicated as green squares in Figure \@ref(fig:dominio)a. They have higher spatial coverage than the CSWS and a sampling frequency of 10 minutes in most cases. All stations measure temperature, but only 395 stations provide humidity, 422 provide pressure, and 605 provide wind information. 
Observation errors used to assimilate these observations are the same as for the CSWS (see Table \@ref(tab:table-obs)).

@@

To evaluate the performance of the ensemble-based DA system presented in this article, the following observational datasets were used:

@@

To investigate the impact of different observations upon the analysis, four DA experiments were performed using different observation sets (Table \@ref(tab:table-exp)). The CONV experiment uses only conventional observations from PREPBUFR. In a second experiment, referred to as AWS, all the observations included in CONV are assimilated plus the 10-minute frequency surface observations from AWS. In the third experiment, referred to as SATWND, the observations from the AWS experiment along with the satellite-derived winds are assimilated. Finally, a fourth experiment referred to as RAD assimilates all available clear-sky radiances from sensors onboard polar orbiting satellites as described in section \@ref(sat).

@@

To evaluate the statistical consistency of the uncertainty quantification in the ensemble system the Reduced Centered Random Variable (RCRV, @candille2007) is used which is defined as:

@@

where $x_o$ is the assimilated observation and its error $\sigma_o$, the ensemble mean of the analysis in observational space $m$, and the standard deviation $\sigma$ of the ensemble. The $RCRV$ is the ratio of the distance between the observations and the forecast and its expected standard deviation assuming the statistical independence between the forecast error (estimated from the ensemble spread) and the observation error. The average of $RCRV$ computed over all the analysis cycles represents the bias of the ensemble mean with respect to the observations normalized by the estimated uncertainty:

@@

For the comparison of the first-guess precipitation with the IMERG precipitation estimates, the Fractions Skill Score (FSS, @roberts2008) is computed for different neighborhood length scales and thresholds:

@@

where $P_{oi}$ is the fraction of grid points in the $i-th$ sampling area in which the observed accumulated precipitation is greater than a specified threshold. Following @roberts2020, $P_{xi}$ is calculated from the ensemble probability precipitation over the same threshold in each grid point by averaging over the $i-th$ sampling area. 
The FSS was computed from the accumulated precipitation over 6 hr rolling windows by adding the 1-hr accumulated precipitation forecasts over 6 consecutive assimilation cycles.

@@

All the experiments were performed at the National Center for Atmospheric Research (NCAR) supercomputer Cheyenne [@Cheyenne2019]. All the analyses in this paper were conducted using the R programming language [@rcoreteam2020], using data.table [@dowle2020] and metR [@campitelli2020] packages.
All graphics are made using ggplot2 [@wickham2009] and the paper was rendered using knitr and rmarkdown [@xie2015; @allaire2019].

@@

To investigate the ability of the first-guess ensemble mean to fit the observations taking into account the uncertainties of the forecast and the observations, the $mean RCRV$ and the $sd RCRV$ is calculated for the RAD experiment. As this experiment assimilates all types of observations used in this work, it is possible to analyze the consistency of the ensemble by comparing it with each type of observation. Figure \@ref(fig:rcrv-sfc) shows the $sd RCRV$ for surface observations box-averaged to a 2.5° grid. The $sd RCRV$ for wind observations (Figure \@ref(fig:rcrv-sfc)a) is close to 1 suggesting a good agreement between the ensemble spread, the forecast error, and the observation error. For the temperature (Figure \@ref(fig:rcrv-sfc)b), the results are similar except that for some areas in the west of the domain the $sd RCRV$ can be as high as 4.5. These higher values of $sdRCRV$ can be associated with systematic errors arising from high differences between the model surface and the observations. Small scale circulations associated with the complex terrain and not well resolved by the model can also contribute to increase the distance between the forecast and the observations. These aspects are usually not captured by the ensemble spread unless a well tuned space dependent inflation scheme is used thus leading to greater sdRCRV values.

Figure \@ref(fig:rcrv-profile) shows the mean and standard deviation of the RCRV for the upper-air observations. Figures \@ref(fig:rcrv-profile)a-b show the RCRV statistics for soundings (ADPUPA) and aircraft (AIRCAR and AIRCFT). Both ADPUPA and AIRCFT show a generally good agreement between the ensemble spread and the observation error. As sounding observations and their associated errors are known to be reliable, this result indicates that the ensemble has an appropriate spread. AIRCAR presents an irregular profile with $sd RCRV$ values that suggest that the error for this type of observation is overestimated. ADPUPA and AIRCAR present a $mean RCRV$ profile near zero at middle and upper levels.

Satellite-derived winds observations varies in number depending on the satellite and the level. In Figure \@ref(fig:rcrv-profile)c only the $RCRV$ calculated with at least 100 observations for each satellite and level is included. At low levels, where there are not many observations available, the profiles of $mean RCRV$ and $sd RCRV$ show a larger departure from the expected behavior with a negative bias, and a possible overestimation of the observation error. Wind estimations derived from water vapor channels are abundant above 500 hPa where their bias is close to zero. The only exception are the EUMETSAT observations which contribute very little in the region.

@@

This section presents the impact of assimilating different observation types on variables which are particularly relevant for the occurrence of deep moist convection. The analysis is performed over a smaller domain (red box in Figure \@ref(fig:dominio)a) to focus on the region most directly affected by the MCS. Figures \@ref(fig:TQ-diff)a-c show the analysis difference between experiments in the spatially averaged vertical profile of temperature. By averaging the differences between two experiments the systematic impact produced by different observing systems on the analyzed state can be isolated. During the first day, the assimilation of AWS observations results in a colder PBL. This cooling effect has a clear diurnal cycle, being stronger during nighttime (Figure \@ref(fig:TQ-diff)a). During the second day of the experiment, the impact of AWS observations extends into the middle and upper troposphere coinciding with the mature stage of the MCS. The warm difference shown in AWS-CONV between 500 and 200 hPa is produced by the development of stronger convection in AWS compared to CONV. This is a good example of how low-level information provided by surface weather stations can rapidly spread into the troposphere in the presence of deep moist convection. Although the mid-to-upper circulation can have an important impact on the organization and evolution of the MCS over the region, the satellite-derived winds did not have an appreciable impact on the mean temperature and humidity (Figure \@ref(fig:TQ-diff)b-e), possibly due to the large observation errors used for the assimilation.
During the first day of the experiment, the assimilation of radiances produces a warming effect in the PBL which partially compensates for the cooling effect of AWS observations (Figure \@ref(fig:TQ-diff)c). No clear systematic impact is found above the PBL during this period. During the second day, the impact of radiance observations is found through the troposphere with a distribution that is similar to the impact found in the AWS experiment but with the opposite sign.

Comparing the specific humidity in the experiments (Figures \@ref(fig:TQ-diff)d-f), the impact of assimilating AWS with fine spatial and temporal resolution is most substantial at low levels (Figure \@ref(fig:TQ-diff)d). The PBL in the AWS experiment is consistently moister than in the CONV experiment particularly at nighttime. The increase in low-level moisture by a denser surface network is consistent with previously reported dry biases in the WRF model over the region [@casaretto2022, @matsudo2021, @ruiz2010]. The moistening of the PBL is mainly driven by the covariance between temperature and specific humidity within the PBL. In the experiment and over the center of the domain, this covariance remains negative, increasing low-level moisture as the observations introduce negative temperature corrections. As for the temperature, the systematic impact of satellite-derived winds on moisture is small (Figure \@ref(fig:TQ-diff)e). Figure \@ref(fig:TQ-diff)f shows that radiances reduce low-middle level moisture during the first day of the experiment. The drying effect extends to lower-middle levels during the second day of the experiment coinciding with the development of the MCS between 00 and 12 UTC Nov 22.

@@

IOP 7 (Figures \@ref(fig:soundings)a-d) provides a good characterization of the pre-convective environment during the first day of our experiments. The area where the observations were taken was characterized by mostly clear skies and a low-level northerly flow associated with warm and moist advection. In general, the experiments show a similar RMSE and bias for all the variables. AWS observations were able to reduce the RMSE for temperature and dew point temperature in the PBL and reduce a small dry bias. However, in this region (Figure \@ref(fig:dominio)b) and for this period, AWS increments (Figure \@ref(fig:UV-diff)d) degrades the zonal wind between 7 and 12 km increasing the bias and RMSE (Figure \@ref(fig:soundings)c).
